Predicting Infection Outbreak Dynamics in Toronto Healthcare Institutions: A Bayesian Approach

### 2024-12-15 paper/paper.qmd
Update paper.qmd

---
title: "Predicting Infection Outbreak Dynamics in Toronto Healthcare Institutions: A Bayesian Approach"
subtitle: "Respiratory Outbreaks Are More Likely in Long-Term Care Homes and Retirement Homes, Driven by Seasonal Trends and Causative Agents Like COVID-19 and Influenza"
author: "Divya Gupta"
date: "December 3, 2024"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

@@

library(tidyverse)
library(palmerpenguins)
```
# Abstract
Identifying key trends in infection outbreaks in a city can help policymakers implement better reforms to support citizens and prevent outbreaks. This study uses cleaned datasets retrieved from Open Data Toronto to develop a model that can have insightful implications for the City of Toronto policymakers, especially regarding resource allocation for outbreak prevention. This study aims to help policymakers better understand the trends of infection outbreaks in Toronto, so that this data can be used to identify problem areas in the city (such as specific times of year). This study examines the factors influencing healthcare outbreaks in Toronto healthcare institutions, such as: type of infection (respiratory vs. gastroenteric), causative agent of infection, setting and month of infection. This study uses Bayesian logistic regression to identify the factors associated with infectious outbreaks and predict the likelihood of a specific outbreak type depending on the aforementioned variables. The findings show that COVID-19 is a significant causative agent for respiratory infections, while norovirus is a causative agent for enteric outbreaks. Moreover, there is a significant increase in the number of outbreaks in winter months and in long term care homes.

@@

# Data {#sec-data}

@@

# Constructed Variables

@@

# Accuracy and Limitations
The Open Data Toronto portal provided the raw dataset that had some inconsistencies, namely - various fields had missing entries in the “Date Declared Over” tab, which rendered those data points useless as we could not track the duration of the outbreak. This value is essential for this analysis as our variable, Duration, was constructed as the difference between “Date Declared Over” and “Date Outbreak Began.” Further, some data points had no specific entry for “Type of Outbreak”, and had values like “other” which did not provide usable information for this analysis. For ease of analysis, these data entries were removed while constructing the analysis dataset. 
Understanding the Data
Outcome Variable 
Our model predicts the probability of an outbreak being labeled as respiratory or enteric based on its predictors. To that effect, our outcome variable is “Type_of_Outbreak”, with its results being in binary form of either respiratory or enteric.

# Predictor Variables

@@

# Visualising the Data

Figure 1 illustrates the difference between the levels of respiratory and enteric outbreaks in the city. As can be seen, the overwhelming amount of outbreaks reported are respiratory. 
Figure 2 takes this one step further and illustrates the number of outbreaks determined by healthcare settings. This shows that LTHC had an overwhelming majority in the number of outbreaks reported, followed by retirement homes.

Next, we aimed to understand the interaction between different types of infection outbreaks and the setting in which they occurred. This is illustrated in Figure 3 which is a stacked bar chart. Key insights included the fact that respiratory infections dominated across all types of healthcare settings. While enteric outbreaks are less common, they are most prevalent in retirement homes and hospitals-acute care. Psychiatric care units in hospitals and transitional care systems reported no cases of enteric outbreaks in this dataset.
Lastly, we aimed to understand the overall interaction trends between the outbreak duration, the causative agent, and the setting of the outbreak. This is illustrated in Figure 4. 
This series of density plots illustrates the duration of outbreaks as stratified by the causative agents and various healthcare settings included in the dataset. This shows that LTHC and Retirement homes report significantly longer outbreak durations, specifically for causative agents such as norovirus and Influenza. Further, transitional care settings have a short outbreak duration. This may be due to strict infection control practices - this raises a question about outbreak control policies in less-monitored settings such as retirement homes which have reported longer outbreak durations. These differences in durations according to healthcare settings also raises a need for setting-tailored outbreak interventions.

@@

3.1 Model Set-Up

3.1 Model Set-Up
This study aims to build a Bayesian Logistic Regression model using R (R Core Team, 2023) to predict the probability of a respiratory outbreaks in Toronto. Based on our exploratory data analysis, we can see that predictor variables such as causative agent, healthcare setting, and month/duration of the outbreak are key predictors in helping model this. The model is, thus, as follows: 
Yi ~ Bernoulli(pii)
logit(pii) = β0​+β1​⋅Outbreak_Settingi​+β2​⋅Causative_Agent_1i​+β3​⋅Monthi​+ϵi​
Where: 
Pi is a predicted probability of a respiratory outbreak for the ith observation. 
Β0​ is the intercept, that captures the baseline log-odds of a respiratory outbreak. 
Outbreak_Setting helps determine the role of healthcare settings in the determination of an outbreak.
Causative_agent captures the role of the primary pathogen that causes the outbreak. 
Month captures the month in which an outbreak started, aiming to capture any seasonal aspects of outbreaks. 
Lastly, ei is an error term capturing any noise in the model.
Priors
This model incorporates the following priors: 
β∼N(0,2.5)\beta \sim N(0, 2.5) for the coefficients of predictors.
β0∼N(0,5)\beta_0 \sim N(0, 5) for the intercept.
These are weakly informative priors that help constrain extreme values of the variables without being extremely restrictive.
3.2 Model Justification
Since this paper deals with a binary variable such as Respiratory infection outbreaks, logistic Bernoulli regression was used. This is so that prior knowledge and uncertainty (through the priors and error term) can be accommodated and binary outcome variables can be handled effectively. 
The logit link function in the model ensures that the predicted probability of a respiratory outbreak is between 0 and 1. A Bayesian approach was taken to the model to allow full posterior distributions for each predictive parameter. Further, it allows for the inclusion of prior knowledge, which can be helpful in understanding outbreak trends over a long period of time. Further, the priors help regularize parameter estimates and prevent overfitting to certain parameters such as outbreak settings or causative agents. 
The predictors were chosen for this model based on our exploratory data analysis, where they were shown to be significant in understanding the type of outbreak. Certain alternative model choices such as regular logistic regression were considered but not used due to their inability to incorporate prior knowledge. The Bayesian model, overall, includes prior knowledge along with a nuance to the convoluted nature of our chosen predictor variable, making it a robust choice for this analysis. 
3.3 Assumptions and Limitations

There are several key assumptions of this model. It assumes that the outcome variable, Type of Outbreak, is a binary outcome (respiratory or enteric). Further, it assumes that each outbreak is an independent event influenced by its own set of predictive variables. However, this assumption may not hold for outbreaks that occurred in the same healthcare setting, or occurred very closely together, or overlapped in the duration or month variables. This model does not account for these potential interactions. Lastly, the weakly informative priors are assumed to appropriately reflect the healthcare setting due to their unrestrictive flexibility. 
This raises certain limitations to the scope of the model. The model assumes the dataset it was developed on was accurately and consistently reported on by all healthcare institutions. It is possible that the causative agent or the duration of the outbreak, for example, could be misreported which would skew the results. Further, nuanced aspects of healthcare settings such as quality of care provided, healthcare policies, severity of symptoms, and staff-to-patient ratios can vary the severity and occurence of the outbreak but these were not captured by this model. While the month variable was included to capture seasonal trends in outbreaks, it is possible that this may vary year to year. Our dataset only captures information for 2024 and may not be generalizable to a wider range of time without proper adjustment. 
3.4 Model Validation
Validation is critical to ensure the models are reliable and provide accurate predictions. As a part of the model validation process, Posterior Predictive Checks were conducted to assess the fit of the model. Figure 5 shows the Posterior Predictive Check that compares the observed data (illustrated by y) and the predicted data (illustrated by yrep). The observed outcome is overlaid with the densities of the simulated outcome to see how well the model’s prediction matches with the real data. 
There is a close alignment between the two types of data, which shows that the model captures the data distribution effectively. There are slight deviations in the lower probability regions which suggests that there are areas for improvement in the model. This could be tackled in the future by including more predictor variables.

@@

Figure 6 displays the posterior estimates and their corresponding 95% credible intervals for each predictor in the model. The x-axis represents the effect size as log-odds while the y-axis shows the predictors. This figure shows that LTCH have a high number of respiratory outbreaks, given that their credible intervals are above zero. Hospital-Chronic Care, Hospital-Psychiatric, and Retirement Homes also have positive effects but have differing magnitudes and credible intervals. Further, causative agents such as COVID-19, Coronavirus, and Respiratory Syncytial Virus have positive associations with respiratory outbreaks as their credible intervals are above zero. Influenza A and Metapneumovirus also have an association with respiratory outbreaks but have wider intervals, showing uncertainty. On the other hand, norovirus is heavily associated with enteric outbreaks as they have a negative association with respiratory outbreaks. There is a wide variability in the month and its association with different outbreaks, illustrating uncertainty in this predictor. 
Overall, we see that LTCH settings (β=2.3) are most associated with respiratory outbreaks, followed by Retirement homes and Hospital- Chronic care. Further, COVID-19, Coronavirus,
and Respiratory Syncytial Virus are the strongest predictors of respiratory outbreaks for causative agents. Causative agents like COVID-19 (β=3.8) strongly increase the likelihood of respiratory outbreaks while norovirus (β=−7.4) strongly increase the likelihood of enteric outbreaks.

@@

# 5.1 Strengths
The inclusion of an array of predictors allowed the model to account for a high level of variability in the data, and understand the effect of different factors on outbreaks. The Bayesian approach allowed the quantification of categorical variables as well as the inclusion of weakly informative priors that allow the incorporation of prior knowledge from healthcare settings. This allowed for key insights, as discussed in the Results section, that showed that LTCH and Retirement Homes are key healthcare settings to be targeted to tackle respiratory outbreaks.

# 5.2 Limitations
This study has several limitations. The dataset relies on accurate reporting by healthcare institutions, which tend to be overworked and can misreport data. Further, multiple data entries were excluded due to incomplete data points, which skews the robustness of the model. The dataset only incorporated information from 2024 and it is unclear if this can be extrapolated to different years. A wider dataset, with more information ranging a 5 or 10 year period, may allow for a more robust model. While some seasonal trends were seen in the analysis, it is still important to note that these are seasonal trends over the span of only one year, and they can vary significantly. More precise week-by-week analysis may be favored over monthly analysis to understand outbreak trends. Further, our outcome variable is binary - respiratory or enteric - and can overlook the aspect of interaction between various types of infections occurring together.

@@

# Appendix A: Idealized Survey: Methodology

@@

Sampling Approach

@@

Non-Response Handling

@@

Strengths & Limitations

@@

Toronto Infection Outbreak Tracking Survey

@@

>>>>

---
title: "Predicting Infection Outbreak Dynamics in Toronto Healthcare Institutions: A Bayesian Approach"
subtitle: "Respiratory Outbreaks Are More Likely in Long-Term Care Homes and Retirement Homes, Driven by Seasonal Trends and Causative Agents Like COVID-19 and Influenza"
author: "Divya Gupta"
thanks: "Code and data are available at: https://github.com/gdivyagupta/Infections_in_Toronto"
date: today
date-format: long
abstract: Identifying key trends in infection outbreaks in a city can help policymakers implement better reforms to support citizens and prevent outbreaks. This study uses cleaned datasets retrieved from Open Data Toronto to develop a model that can have insightful implications for the City of Toronto policymakers, especially regarding resource allocation for outbreak prevention. This study aims to help policymakers better understand the trends of infection outbreaks in Toronto, so that this data can be used to identify problem areas in the city (such as specific times of year). This study examines the factors influencing healthcare outbreaks in Toronto healthcare institutions, such as: type of infection (respiratory vs. gastroenteric), causative agent of infection, setting and month of infection. This study uses Bayesian logistic regression to identify the factors associated with infectious outbreaks and predict the likelihood of a specific outbreak type depending on the aforementioned variables. The findings show that COVID-19 is a significant causative agent for respiratory infections, while norovirus is a causative agent for enteric outbreaks. Moreover, there is a significant increase in the number of outbreaks in winter months and in long term care homes. 
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
```

@@

# Data {#paper-data-section}

@@

## Constructed Variables

@@

## Accuracy and Limitations
The Open Data Toronto portal provided the raw dataset that had some inconsistencies, namely - various fields had missing entries in the “Date Declared Over” tab, which rendered those data points useless as we could not track the duration of the outbreak. This value is essential for this analysis as our variable, Duration, was constructed as the difference between “Date Declared Over” and “Date Outbreak Began.” Further, some data points had no specific entry for “Type of Outbreak”, and had values like “other” which did not provide usable information for this analysis. For ease of analysis, these data entries were removed while constructing the analysis dataset. 
Understanding the Data
Outcome Variable 
Our model predicts the probability of an outbreak being labeled as respiratory or enteric based on its predictors. To that effect, our outcome variable is “Type_of_Outbreak”, with its results being in binary form of either respiratory or enteric.

## Predictor Variables

@@

## Visualising the Data

Figure 1 illustrates the difference between the levels of respiratory and enteric outbreaks in the city. As can be seen, the overwhelming amount of outbreaks reported are respiratory.

```{r}
cleaned_data |> 
  ggplot(aes(x = `Type of Outbreak`, fill = `Type of Outbreak`)) +
  geom_bar() +
  labs(
    title = "Distribution of Outbreak Types",
    x = "Type of Outbreak",
    y = "Count"
  ) +
  theme_minimal()
```

Figure 2 takes this one step further and illustrates the number of outbreaks determined by healthcare settings. This shows that LTHC had an overwhelming majority in the number of outbreaks reported, followed by retirement homes.

```{r}
cleaned_data |> 
  ggplot(aes(x = fct_infreq(`Outbreak Setting`), fill = `Outbreak Setting`)) +
  geom_bar() +
  coord_flip() +
  labs(
    title = "Outbreak Counts by Healthcare Setting",
    x = "Healthcare Setting",
    y = "Count"
  ) +
  theme_minimal()
```

Next, we aimed to understand the interaction between different types of infection outbreaks and the setting in which they occurred. This is illustrated in Figure 3. Key insights included the fact that respiratory infections dominated across all types of healthcare settings.

```{r}
cleaned_data |> 
  ggplot(aes(x = `Outbreak Setting`, fill = `Type of Outbreak`)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(
    title = "Proportion of Outbreak Types by Setting",
    x = "Healthcare Setting",
    y = "Proportion"
  ) +
  theme_minimal()
```{r}

While enteric outbreaks are less common, they are most prevalent in retirement homes and hospitals-acute care. Psychiatric care units in hospitals and transitional care systems reported no cases of enteric outbreaks in this dataset.
Lastly, we aimed to understand the overall interaction trends between the outbreak duration, the causative agent, and the setting of the outbreak. This is illustrated in Figure 4.

```{r}
cleaned_data |> 
  ggplot(aes(x = Duration, fill = `Causative Agent-1`, color = `Causative Agent-1`)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ `Outbreak Setting`, scales = "free_y") +
  labs(
    title = "Outbreak Duration by Causative Agent and Setting",
    x = "Outbreak Duration (days)",
    y = "Density",
    fill = "Causative Agent",
    color = "Causative Agent"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10)
  )
```{r}

This series of density plots illustrates the duration of outbreaks as stratified by the causative agents and various healthcare settings included in the dataset. This shows that LTHC and Retirement homes report significantly longer outbreak durations, specifically for causative agents such as norovirus and Influenza. Further, transitional care settings have a short outbreak duration. This may be due to strict infection control practices - this raises a question about outbreak control policies in less-monitored settings such as retirement homes which have reported longer outbreak durations. These differences in durations according to healthcare settings also raises a need for setting-tailored outbreak interventions.

@@

## 3.1 Model Set-Up
This study aims to build a Bayesian Logistic Regression model using R (R Core Team, 2023) to predict the probability of a respiratory outbreaks in Toronto. Based on our exploratory data analysis, we can see that predictor variables such as causative agent, healthcare setting, and month/duration of the outbreak are key predictors in helping model this. The model is, thus, as follows:

```{=tex}
\[
y_i \mid \pi_i \sim \text{Bern}(\pi_i)
\]

\[
\text{logit}(\pi_i) = \beta_0 + \beta_1 \times \text{Outbreak\_Setting}_i + \beta_2 \times \text{Causative\_Agent\_1}_i + \beta_3 \times \text{Month}_i + \epsilon_i
\]

\[
\beta_0 \sim \text{Normal}(0, 5)
\]
\[
\beta_1 \sim \text{Normal}(0, 2.5)
\]
\[
\beta_2 \sim \text{Normal}(0, 2.5)
\]
\[
\beta_3 \sim \text{Normal}(0, 2.5)
\]
```

Where: 
𝜋𝑖 is a predicted probability of a respiratory outbreak for the 𝑖th observation. 
β0 is the intercept that captures the baseline log-odds of a respiratory outbreak. 
β1, β2, and β3 are coefficients for the categorical predictors.
Outbreak_Setting𝑖​ helps determine the role of healthcare settings in the determination of an outbreak.
Causative_Agent𝑖​ captures the role of the primary pathogen that causes the outbreak. 
Month𝑖​ captures the month in which an outbreak started, aiming to capture any seasonal aspects of outbreaks. 
Lastly,  ϵ𝑖​ is an error term capturing any noise in the model.

## Priors
This model incorporates the following priors: 
β∼N(0,2.5)\beta \sim N(0, 2.5) for the coefficients of predictors.
β0∼N(0,5)\beta_0 \sim N(0, 5) for the intercept.
These are weakly informative priors that help constrain extreme values of the variables without being extremely restrictive.

## 3.2 Model Justification
Since this paper deals with a binary variable such as Respiratory infection outbreaks, logistic Bernoulli regression was used. This is so that prior knowledge and uncertainty (through the priors and error term) can be accommodated and binary outcome variables can be handled effectively. 
The logit link function in the model ensures that the predicted probability of a respiratory outbreak is between 0 and 1. A Bayesian approach was taken to the model to allow full posterior distributions for each predictive parameter. Further, it allows for the inclusion of prior knowledge, which can be helpful in understanding outbreak trends over a long period of time. Further, the priors help regularize parameter estimates and prevent overfitting to certain parameters such as outbreak settings or causative agents. 
The predictors were chosen for this model based on our exploratory data analysis, where they were shown to be significant in understanding the type of outbreak. Certain alternative model choices such as regular logistic regression were considered but not used due to their inability to incorporate prior knowledge. The Bayesian model, overall, includes prior knowledge along with a nuance to the convoluted nature of our chosen predictor variable, making it a robust choice for this analysis.

## 3.3 Assumptions and Limitations

There are several key assumptions of this model. It assumes that the outcome variable, Type of Outbreak, is a binary outcome (respiratory or enteric). Further, it assumes that each outbreak is an independent event influenced by its own set of predictive variables. However, this assumption may not hold for outbreaks that occurred in the same healthcare setting, or occurred very closely together, or overlapped in the duration or month variables. This model does not account for these potential interactions. Lastly, the weakly informative priors are assumed to appropriately reflect the healthcare setting due to their unrestrictive flexibility. 
This raises certain limitations to the scope of the model. The model assumes the dataset it was developed on was accurately and consistently reported on by all healthcare institutions. It is possible that the causative agent or the duration of the outbreak, for example, could be misreported which would skew the results. Further, nuanced aspects of healthcare settings such as quality of care provided, healthcare policies, severity of symptoms, and staff-to-patient ratios can vary the severity and occurence of the outbreak but these were not captured by this model. While the month variable was included to capture seasonal trends in outbreaks, it is possible that this may vary year to year. Our dataset only captures information for 2024 and may not be generalizable to a wider range of time without proper adjustment.

## 3.4 Model Validation
Validation is critical to ensure the models are reliable and provide accurate predictions. As a part of the model validation process, Posterior Predictive Checks were conducted to assess the fit of the model. Figure 5 shows the Posterior Predictive Check that compares the observed data (illustrated by y) and the predicted data (illustrated by yrep). The observed outcome is overlaid with the densities of the simulated outcome to see how well the model’s prediction matches with the real data. 
There is a close alignment between the two types of data, which shows that the model captures the data distribution effectively. There are slight deviations in the lower probability regions which suggests that there are areas for improvement in the model. This could be tackled in the future by including more predictor variables.

```{r}
# Default posterior predictive check (histogram)
pp_check(logistic_model) +
  labs(
    title = "Posterior Predictive Check: Observed vs. Predicted Outcomes",
    x = "Outcome (Type of Outbreak: Enteric = 0, Respiratory = 1)",
    y = "Density"
  ) +
  theme_minimal()

# Overlay observed and predicted densities (use density comparison)
pp_check(logistic_model, overlay = TRUE) +
  labs(
    title = "Posterior Predictive Check: Density Overlay",
    x = "Probability of Respiratory Outbreak",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Legend")) +
  scale_color_manual(
    values = c("y" = "black", "yrep" = "lightblue"),
    labels = c("Observed Data (y)", "Predicted Data (yrep)")
  )

# Visualisations with the model
logistic_model |> posterior_summary()
```{r}

@@

Figure 6 displays the posterior estimates and their corresponding 95% credible intervals for each predictor in the model. The x-axis represents the effect size as log-odds while the y-axis shows the predictors. This figure shows that LTCH have a high number of respiratory outbreaks, given that their credible intervals are above zero. Hospital-Chronic Care, Hospital-Psychiatric, and Retirement Homes also have positive effects but have differing magnitudes and credible intervals. Further, causative agents such as COVID-19, Coronavirus, and Respiratory Syncytial Virus have positive associations with respiratory outbreaks as their credible intervals are above zero. Influenza A and Metapneumovirus also have an association with respiratory outbreaks but have wider intervals, showing uncertainty. On the other hand, norovirus is heavily associated with enteric outbreaks as they have a negative association with respiratory outbreaks. There is a wide variability in the month and its association with different outbreaks, illustrating uncertainty in this predictor.

```{r}
# Plot credible intervals with axis labels and title
plot(logistic_model, type = "intervals") +
  labs(
    title = "Posterior Estimates and 95% Credible Intervals",
    x = "Effect Size (Log-Odds)",
    y = "Predictors"
  ) +
  theme_minimal()

# Generate predicted values
predicted <- ggpredict(logistic_model, terms = c("Outbreak_Setting", "Causative_Agent_1"))

# Plot with sufficient colors
plot(predicted) +
  scale_color_manual(values = scales::hue_pal()(20)) +  # Generate a palette with at least 20 colors
  labs(
    title = "Predicted Probabilities by Outbreak Setting and Causative Agent",
    x = "Outbreak Setting",
    y = "Predicted Probability",
    color = "Causative Agent"
  ) +
  theme_minimal()
```{r}

Overall, we see that LTCH settings (β=2.3) are most associated with respiratory outbreaks, followed by Retirement homes and Hospital- Chronic care. Further, COVID-19, Coronavirus,
and Respiratory Syncytial Virus are the strongest predictors of respiratory outbreaks for causative agents. Causative agents like COVID-19 (β=3.8) strongly increase the likelihood of respiratory outbreaks while norovirus (β=−7.4) strongly increase the likelihood of enteric outbreaks.

@@

## 5.1 Strengths
The inclusion of an array of predictors allowed the model to account for a high level of variability in the data, and understand the effect of different factors on outbreaks. The Bayesian approach allowed the quantification of categorical variables as well as the inclusion of weakly informative priors that allow the incorporation of prior knowledge from healthcare settings. This allowed for key insights, as discussed in the Results section, that showed that LTCH and Retirement Homes are key healthcare settings to be targeted to tackle respiratory outbreaks.

## 5.2 Limitations
This study has several limitations. The dataset relies on accurate reporting by healthcare institutions, which tend to be overworked and can misreport data. Further, multiple data entries were excluded due to incomplete data points, which skews the robustness of the model. The dataset only incorporated information from 2024 and it is unclear if this can be extrapolated to different years. A wider dataset, with more information ranging a 5 or 10 year period, may allow for a more robust model. While some seasonal trends were seen in the analysis, it is still important to note that these are seasonal trends over the span of only one year, and they can vary significantly. More precise week-by-week analysis may be favored over monthly analysis to understand outbreak trends. Further, our outcome variable is binary - respiratory or enteric - and can overlook the aspect of interaction between various types of infections occurring together.

# Conclusion 
This study provides valuable insights into the dynamics of infection outbreaks in Toronto's healthcare institutions, focusing on respiratory and enteric outbreaks. By employing a Bayesian logistic regression model, we identified significant predictors of outbreak types, including healthcare setting, causative agents, and seasonal trends. These findings shed light on the vulnerability of the Toronto healthcare system, specifically of long-term care homes and retirement homes. During winter months, these spaces are more likely to experience respiratory outbreaks caused by COVID-19 or influenza.
These key findings allow policymakers to develop targeted interventions for at-risk healthcare establishments in the city. This can be in the form of data-informed infection control practices, seasonal changes in resource allocation, and better information distribution. Further work is needed to enhance this study’s findings and inform policy development by pooling data from more than just 2024 and analysing decades-long trends in infection outbreaks in the city. Moreover, the study can be further built upon by developing a model to locate geographic hotspots for infections in the city and developing district-based policies to combat this issue.

\newpage

@@

## Appendix A: Idealized Survey: Methodology

@@

## Sampling Approach

@@

## Non-Response Handling

@@

## Strengths & Limitations

@@

## Toronto Infection Outbreak Tracking Survey

@@

\newpage
